systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2004-11-16 paper/trunk/systemfit.tex
merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@567 591db2bd-e0a0-4d22-8c19-2192dbb7da31

If no cross-equation restrictions are imposed, 
the parameter estimates and the variance-covariance matrix 
of the parameters are equal to the OLS estimates%
\footnote{The variance-covariance matrices of the 
estimated parameters might differ, 
if the WLS estimation is based on a different formula 
to calculate the (diagonal) variance-covariance 
matrix of the residuals (see section ????).}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Residual variance-covariance matrix}

Since the true residuals of the estimated equations are generally not known,
the true covariance matrix of the residuals cannot be determined.
Thus, the this covariance matrix must be calculated from the
\emph{estimated} residuals.

The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
\left[ \sigma \right]$) 
is often calculated by
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
   \label{eq:rcov0}
\end{equation}
where $T$ is the number of observations in each equation.

However, this estimator is biased, because it is not corrected for the
degrees of freedom. 
The usual single-equation procedure to correct for degrees of freedom 
cannot always be applied, because the number of regressors in each equation 
might differ.
Two alternative approaches to calculate the residual variance-covariance 
matrix are
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
   \label{eq:rcov1}
\end{equation}
and
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { T - \max \left( K_i , K_j \right) }
\end{equation}
where $K_i$ and $K_j$ are the number of regressors in equation
$i$ and $j$, respectively.

@@

A further approach to obtain the estimated residual variance-covariance
matrix is
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] }
   \label{eq:rcov2}
\end{equation} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, "it is doubtful whether [this formula] is really superior to 
[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.

@@

The other mandatory argument \code{eqns} is a list of the equations 
to estimate. 
Each equation is a standard formula in \proglang{R}.
It starts with a dependent variable on the left hand side.
After a tilde ($~$) the regressors are listed%
\footnote{For Details see the \proglang{R} help files to \code{formula}}.

@@

\section{CVS Revision Log}
\begin{verbatim}
$Log$
Revision 1.7  2004/11/16 05:34:47  hamannj
question about nlsystemfit code section (example)

>>>>

If no cross-equation restrictions are imposed, 
the parameter estimates and the variance-covariance matrix 
of the parameters are equal to the OLS estimates%
\footnote{The variance-covariance matrices of the 
estimated parameters might differ, 
if the WLS estimation is based on a different formula 
to calculate the (diagonal) variance-covariance 
matrix of the residuals (see section \ref{sec:residcov}).}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Residual variance-covariance matrix}\label{sec:residcov}

Since the true residuals of the estimated equations are generally not known,
the true covariance matrix of the residuals cannot be determined.
Thus, this covariance matrix must be calculated from the
\emph{estimated} residuals.

The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
\left[ \widehat{\sigma}_{ij} \right]$) 
is often calculated by
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
   \label{eq:rcov0}
\end{equation}
where $T$ is the number of observations in each equation.

However, this estimator is biased, because it is not corrected for 
degrees of freedom. 
The usual single-equation procedure to correct for degrees of freedom 
cannot always be applied, because the number of regressors in each equation 
might differ.
Two alternative approaches to calculate the residual variance-covariance 
matrix are
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
   \label{eq:rcov1}
\end{equation}
and
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { T - \max \left( K_i , K_j \right) }
\end{equation}
where $K_i$ and $K_j$ are the number of regressors in equation
$i$ and $j$, respectively.

@@

A further approach to obtain the estimated residual variance-covariance
matrix is \citep[p. 309]{zellner62c}
\begin{eqnarray}
   \widehat{\sigma}_{ij} & = & 
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
   & = &
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
   \label{eq:rcov2}
\end{eqnarray} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, "it is doubtful whether [this formula] is really superior to 
[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.

@@

The other mandatory argument \code{eqns} is a list of the equations 
to estimate. 
Each equation is a standard formula in \proglang{R}.
It starts with a dependent variable on the left hand side.
After a tilde ($\sim$) the regressors are listed%
\footnote{For Details see the \proglang{R} help files to \code{formula}}.

@@

\section{CVS Revision Log}
\begin{verbatim}
$Log$
Revision 1.8  2004/11/16 08:55:53  henningsena
merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.

Revision 1.7  2004/11/16 05:34:47  hamannj
question about nlsystemfit code section (example)
